---
title: "PART 2: Signature profile visualization and analysis"
author: ["Shixiang Wang, Ziyu Tao, Tao Wu, Xue-Song Liu (Corresponding author)"]
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    toc_depth: 3
bibliography: ref.bib
link-citations: yes
---


```{r setup-02, include=FALSE}
library(knitr)
library(rmdformats)
## Global options
options(max.print = "75")
opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  prompt = FALSE,
  tidy = "styler",
  comment = NA,
  dpi = 300,
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
opts_knit$set(width = 75)
Sys.setenv("LANGUAGE" = "EN")
```

In this part, we would like to visualize and analyze all tumor samples as a whole.

## Signature profile visualization

```{r}
library(tidyverse)
library(sigminer)

pcawg_sigs <- readRDS(file = "../data/pcawg_cn_sigs_CN176_signature.rds")
```

Show signature profile with relative contribution value for each component.

```{r, fig.height=10, fig.width=14}
show_sig_profile(
  pcawg_sigs,
  mode = "copynumber",
  method = "X",
  style = "cosmic"
)
```

Show signature profile with absolute contribution value (estimated segment count) for each component.

```{r, fig.height=10, fig.width=14}
show_sig_profile(
  pcawg_sigs,
  mode = "copynumber",
  method = "X",
  style = "cosmic",
  normalize = "raw"
)
```

Show signature activity profile.

```{r fig.width=14, fig.height=7}
show_sig_exposure(
  pcawg_sigs,
  style = "cosmic",
  rm_space = TRUE,
  palette = sigminer:::letter_colors
)
```

## Signature analysis

### Signature similarity

Here we explore the similarity between copy number signatures identified in PCAWG data. High cosine similarity will cause some issues in signature assignment and activity attribution [@mauraPracticalGuideMutational2019].

```{r}
sim <- get_sig_similarity(pcawg_sigs, pcawg_sigs)
```

```{r, fig.width=5.5, fig.height=5}
pheatmap::pheatmap(sim$similarity, cluster_cols = F, cluster_rows = F, display_numbers = TRUE)
```

Most of signature-pairs have similarity around `0.1`.

### Signature activity distribution

Transform data to long format.

```{r}
df_abs <- get_sig_exposure(pcawg_sigs) %>% 
  tidyr::pivot_longer(cols = starts_with("Sig"), names_to = "sig", values_to = "activity") %>% 
  dplyr::mutate(activity = log10(activity + 1),
                sig = factor(sig, levels = paste0("Sig", 1:11)))

df_rel <- get_sig_exposure(pcawg_sigs, type = "relative", rel_threshold = 0) %>% 
  tidyr::pivot_longer(cols = starts_with("Sig"), names_to = "sig", values_to = "activity") %>% 
  dplyr::mutate(
    sig = factor(sig, levels = paste0("Sig", 1:11))
  )
```

Plot with function in `sigminer`.

```{r}
show_group_distribution(
  df_abs,
  gvar = "sig",
  dvar = "activity",
  order_by_fun = FALSE,
  g_angle = 90,
  ylab = "log10(activity+1)"
)
```

```{r}
show_group_distribution(
  df_rel,
  gvar = "sig",
  dvar = "activity",
  order_by_fun = FALSE,
  g_angle = 90,
  ylab = "relative activity"
)
```

## Copy number profile and signature activity

### Copy number profile for representative samples

Here we will draw copy number profile of 2 representative samples for each signature.

```{r}
get_sample_from_sig <- function(dt, sig) {
  res = head(dt[order(dt[[sig]], decreasing = TRUE)], 2L)
  res
}
```

Read `CopyNumber` object for PCAWG data.

```{r}
pcawg_cn_obj <- readRDS("../data/pcawg_cn_obj.rds")
samp_summary <- pcawg_cn_obj@summary.per.sample
rel_activity <- get_sig_exposure(pcawg_sigs, type = "relative", rel_threshold = 0)
rel_activity <- rel_activity[, lapply(.SD, function(x) {if (is.numeric(x)) round(x, 3) else x})]
```


```{r, fig.width=12, fig.height=7}
dir.create("../output/enrich_samples/", showWarnings = FALSE)
for (i in paste0("Sig", 1:11)) {
  cat(paste0("Most enriched in ", i, "\n"))
  s = get_sample_from_sig(rel_activity, i)
  print(s)
  plist = show_cn_profile(pcawg_cn_obj, samples = s$sample,
                          show_title = TRUE,
                          return_plotlist = TRUE)
  plist = purrr::map2(plist, s$sample, function(p, x) {
    s = samp_summary[sample == x]
    text = paste0(
      "n_of_seg:", s$n_of_seg, "\n",
      "n_of_amp:", s$n_of_amp, "\n",
      "n_of_del:", s$n_of_del, "\n"
    )
    p <- p + annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1,
                      label = text)
    p
  })
  p = cowplot::plot_grid(plotlist = plist, nrow = 2)
  print(p)
  ggplot2::ggsave(file.path("../output/enrich_samples/", paste0(i, ".pdf")),
                    plot = p, width = 12, height = 6)
}
```

### Ideograph for copy number profile reconstruction

> TODO: select one/two samples and draw copy number profile -> catalog profile -> signature relative activity. **Similar to Nature paper**.


```{r, include=FALSE}
# Clear environment
rm(list = ls())
```

