---
title: "PART 4: Cancer subtyping and prognosis analysis"
author: ["Shixiang Wang, Ziyu Tao, Tao Wu, Xue-Song Liu (Corresponding author)"]
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    toc_depth: 3
bibliography: ref.bib
link-citations: yes
---


```{r setup-04, include=FALSE}
library(knitr)
library(rmdformats)
## Global options
options(max.print = "75")
opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  prompt = FALSE,
  tidy = "styler",
  comment = NA,
  dpi = 300,
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
opts_knit$set(width = 75)
Sys.setenv("LANGUAGE" = "EN")
```

```{r include=FALSE}
ReTidying = TRUE
```


In this part, we would like to cluster all tumors based on CN signatures and use the clusters to explore Pan-Cancer heterogeneity.

## Data integration

Many raw data have been processed by tools (or collected) and cleaned. This part I mainly describe how to combine them to construct the integrated dataset.

> If you care about the data (source) shown below, please check the last section of **PART 0**.

```{r eval=ReTidying}
library(pacman)
p_load(
  tidyverse,
  data.table,
  sigminer,
  IDConverter
)
```

Load pre-processed data.

```{r eval=ReTidying}
pcawg_cn_obj <- readRDS("../data/pcawg_cn_obj.rds")
pcawg_cn_sig <- readRDS("../data/pcawg_cn_sigs_CN176_signature.rds")
pcawg_samp_info_sp <- readRDS("../data/pcawg_samp_info_sp.rds")
samp_summary <- pcawg_cn_obj@summary.per.sample
samp_summary$n_loh <- NULL
```


Check if all samples are recorded.

```{r eval=ReTidying}
all(pcawg_cn_obj@summary.per.sample$sample %in% pcawg_samp_info_sp$pcawg_specimen_histology_August2016_v9$icgc_specimen_id)
```

```{r eval=ReTidying}
pcawg_samp_info_sp <- lapply(pcawg_samp_info_sp, function(x) {
  colnames(x)[1] <- "sample"
  x
})
pcawg_samp_info <- purrr::reduce(pcawg_samp_info_sp, dplyr::full_join, by = "sample")
pcawg_samp_info <- pcawg_samp_info %>% dplyr::filter(sample %in% pcawg_cn_obj@summary.per.sample$sample)

rm(pcawg_samp_info_sp)
```

Refer to @pcawgmutationalsignaturesworkinggroupRepertoireMutationalSignatures2020, some cancer types with small sample size should be combined.

```{r eval=ReTidying}
table(pcawg_samp_info$histology_abbreviation)
```

```{r eval=ReTidying}
pcawg_samp_info <- pcawg_samp_info %>% 
  mutate(
    cancer_type = case_when(
      startsWith(histology_abbreviation, "Breast") ~ "Breast",
      startsWith(histology_abbreviation, "Biliary-AdenoCA") ~ "Biliary-AdenoCA",
      histology_abbreviation %in% c("Bone-Epith", "Bone-Benign") ~ "Bone-Other",
      startsWith(histology_abbreviation, "Cervix") ~ "Cervix",
      histology_abbreviation %in% c("Myeloid-AML, Myeloid-MDS", "Myeloid-AML, Myeloid-MPN", "Myeloid-MDS", "Myeloid-MPN") ~ "Myeloid-MDS/MPN",
      TRUE ~ histology_abbreviation
    )
  )
```

Select important features.

```{r eval=ReTidying}
pcawg_samp_info <- pcawg_samp_info %>% 
  dplyr::select(c("sample", "_PATIENT", "donor_sex", "donor_age_at_diagnosis",
                      "donor_survival_time", "donor_vital_status", "first_therapy_type",
                      "tobacco_smoking_history_indicator", "alcohol_history",
                      "dcc_project_code", "dcc_specimen_type",
                      "histology_abbreviation", "cancer_type",
                      "tumour_stage",
                      "level_of_cellularity"))
colnames(pcawg_samp_info)[2] <- "donor_id"

# Remove duplicated samples
pcawg_samp_info <- pcawg_samp_info[!duplicated(pcawg_samp_info$sample), ]
```

Get signature activities.

```{r eval=ReTidying}
pcawg_activity <- readRDS("../data/pcawg_cn_sigs_CN176_activity.rds")

expo_abs <- pcawg_activity$abs_activity[, lapply(.SD, function(x) if (is.numeric(x)) round(x, digits = 3) else x)]
colnames(expo_abs) <- c("sample", paste0("Abs_", colnames(expo_abs)[-1]))


expo_rel <- pcawg_activity$rel_activity[, lapply(.SD, function(x) if (is.numeric(x)) round(x, digits = 3) else x)]
colnames(expo_rel) <- c("sample", paste0("Rel_", colnames(expo_rel)[-1]))

pcawg_activity2 <- merge(expo_abs, expo_rel, by = "sample")
pcawg_activity2$similarity <- pcawg_activity$similarity
pcawg_activity <- pcawg_activity2
pcawg_activity$keep <- pcawg_activity$similarity >= 0.75

rm(pcawg_activity2, expo_abs, expo_rel)
```

Read tumor purity & ploidy & WGD status.

```{r eval=ReTidying}
dt_pp <- fread("../data/PCAWG/consensus.20170217.purity.ploidy_sp")
dt_pp <- dt_pp[, c("samplename", "purity", "ploidy", "wgd_status")]
colnames(dt_pp)[1] <- "sample" 
```

Calculate fusion events in all samples.

```{r eval=ReTidying}
fusion <- fread("../data/PCAWG/pcawg3_fusions_PKU_EBI.gene_centric.sp.xena")
fusion <- as.matrix(fusion[, -1])
fusion <- apply(fusion, 2, sum)
fusion <- dplyr::tibble(
  sample = names(fusion),
  n_fusion = as.numeric(fusion)
)
```

Load HRD status in all samples.

```{r eval=ReTidying}
HRD <- readRDS("../data/PCAWG/pcawg_CHORD.rds")
HRD <- HRD %>% 
  dplyr::select(icgc_specimen_id, is_hrd, genotype_monoall) %>% 
  set_names(c("sample", "hrd_status", "hrd_genotype")) %>% 
  unique()
```

```{r, include=FALSE, eval=FALSE}
#HRD <- readRDS("../data/PCAWG/PCAWG_HRDindex.RDS")
# HRD <- dplyr::tibble(
#   sample = names(HRD),
#   HRDindex = as.numeric(HRD)
# )
```


Read LOH data.

```{r eval=ReTidying}
LOH <- readRDS("../data/pcawg_loh.rds")
```


Read Chromothripsis data and clean it.

```{r eval=ReTidying}
# http://compbio.med.harvard.edu/chromothripsis/
chromothripsis <- fread("../data/PCAWG/PCAWG-ShatterSeek.txt.gz")
chromothripsis <- chromothripsis[, c("icgc_donor_id", "type_chromothripsis", "chromo_label")]

# Check label
table(chromothripsis$chromo_label)
any(is.na(chromothripsis$chromo_label))

# Summarize
chromothripsis_summary <- chromothripsis[
  , .(n_chromo = sum(chromo_label != "No"),
      chromo_type = paste0(na.omit(type_chromothripsis), collapse = "/")), by = icgc_donor_id]

chromothripsis_summary <- merge(chromothripsis_summary, pcawg_full[, c("icgc_donor_id", "icgc_specimen_id")],
      by = "icgc_donor_id") %>% unique()
chromothripsis_summary$icgc_donor_id <- NULL
colnames(chromothripsis_summary)[3] <- "sample"
rm(chromothripsis)
```

Read AmpliconArchitect results.

```{r eval=ReTidying}
aa1 <- readxl::read_excel("../data/PCAWG/ecDNA.xlsx", skip = 1)
aa2 <- readxl::read_excel("../data/PCAWG/ecDNA.xlsx", sheet = 3)

table(aa1$amplicon_classification)
table(aa2$sample_classification)
all(aa1$sample_barcode %in% aa2$sample_barcode)

aa_summary <- aa1 %>% 
  dplyr::group_by(sample_barcode) %>% 
  dplyr::summarise(
    n_amplicon_BFB = sum(amplicon_classification == "BFB", na.rm = TRUE),
    n_amplicon_ecDNA = sum(amplicon_classification == "Circular", na.rm = TRUE),
    n_amplicon_HR = sum(amplicon_classification == "Heavily-rearranged", na.rm = TRUE),
    n_amplicon_Linear = sum(amplicon_classification == "Linear", na.rm = TRUE),
    .groups = "drop"
  )

data.table::setDT(aa_summary)

custom_dt <- IDConverter::pcawg_simple[, c("submitted_specimen_id", "icgc_specimen_id")]
custom_dt <- custom_dt[startsWith(submitted_specimen_id, "TCGA")]
custom_dt[, submitted_specimen_id := substr(submitted_specimen_id, 1, 15)]

aa_summary[
  , sample := ifelse(
    startsWith(sample_barcode, "SA"),
    convert_pcawg(sample_barcode, from = "icgc_sample_id", to = "icgc_specimen_id"),
    convert_custom(sample_barcode, from = "submitted_specimen_id", to = "icgc_specimen_id", dt = custom_dt))]

## Filter out samples not in PCAWG
aa_summary <- aa_summary[!is.na(sample)][, sample_barcode := NULL]

rm(aa1, aa2, custom_dt)
```

Read TelomereHunter results.

```{r eval=ReTidying, warning=FALSE}
TC <- readxl::read_excel("../data/PCAWG/TelomereContent.xlsx")
TC <- TC %>% dplyr::select(c("icgc_specimen_id", "tel_content_log2"))
colnames(TC)[1] <- "sample"
data.table::setDT(TC)
```

```{r eval=ReTidying}
apobec <- read_tsv("../data/PCAWG/MAF_Aug31_2016_sorted_A3A_A3B_comparePlus.txt", col_types = cols())
apobec <- apobec %>% 
  dplyr::select(c(1, 5)) %>% 
  data.table::as.data.table() %>% 
  setNames(c("sample", "APOBEC_mutations"))
```


Combine and save result combined data.

```{r eval=ReTidying}
pcawg_samp_info <- purrr::reduce(list(pcawg_samp_info,
                                      pcawg_activity,
                                      samp_summary,
                                      dt_pp,
                                      fusion,
                                      HRD,
                                      LOH,
                                      chromothripsis_summary,
                                      aa_summary,
                                      TC,
                                      apobec), 
                                 dplyr::left_join, by = "sample")

saveRDS(pcawg_samp_info, file = "../data/pcawg_sample_tidy_info.rds")

## Remove all objects
rm(list = ls())
```


## Clustering with signature activity



## Heatmap for clustering and genotype/phenotype features

## Exploration of patients' prognosis difference across clusters
