---
title: "PART 5: Inference of copy number signature etiologiess"
author: ["Shixiang Wang, Ziyu Tao, Tao Wu, Xue-Song Liu (Corresponding author)"]
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    toc_depth: 3
bibliography: ref.bib
link-citations: yes
---


```{r setup-05, include=FALSE}
library(knitr)
library(rmdformats)
## Global options
options(max.print = "75")
opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  prompt = FALSE,
  tidy = "styler",
  comment = NA,
  dpi = 300,
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
opts_knit$set(width = 75)
Sys.setenv("LANGUAGE" = "EN")
```

In this part, we would like to explore CN signatures in more details with association analysis and summarize what we get to infer their etiologies.

## Association analysis for copy number signatures

```{r}
pcawg_expo <- readRDS("../data/pcawg_cn_sigs_CN176_activity.rds")
keep_samples <- pcawg_expo$similarity >= 0.75
pcawg_expo$abs_activity <- pcawg_expo$abs_activity[keep_samples]
pcawg_expo$rel_activity <- pcawg_expo$rel_activity[keep_samples]
```

### Signature activity association

For absolute signature activity.

```{r}
show_cor(pcawg_expo$abs_activity[, -1])
```

For relative signature activity.

```{r}
show_cor(pcawg_expo$rel_activity[, -1])
```

### Signature clustering

We can take a try to cluster signatures based on their relative activities.

```{r}
expo_mat_rel <- pcawg_expo$rel_activity %>% 
  tibble::column_to_rownames("sample") %>% 
  as.matrix() %>% 
  t()

hc <- hclust(dist(expo_mat_rel))
```

```{r}
factoextra::fviz_dend(hc, k = 3, rect = T)
```

## Association between copy number signatures (CNS) and other types of signatures (SBS, DBS, ID)

### CNS and SBS

### CNS and DBS

### CNS and ID

## Signature similarity comparison between different types of signatures

> 比较不同 signature type 的平均相似性

## Signature association network construction and visualization

> Try to learn from NC paper

## Association between CNS and genotype/phenotype features

## Association between CNS and patients' prognosis

## Data summary and inference of signature etiologies
